---
title: "Description of Data"
author: "Jonathan Giguere"
date: "10/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = TRUE)
```

```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r, packages, include = F}
library(magrittr)
library(ggplot2)
library(plyr)
library(dplyr)
library(viridis)
library(viridisLite)
```

## Section 1: Introduction

The popularity of smartphones has increased greatly every year since Apple released the Iphone in 2007, with approximately 10 billion [Statista](https://statista.com) smartphones sold worldwide since then. Of these 10 billion, Iphones account for almost 20% of sales with a total revenue of $165 billion in 2018 alone [CEO World Magazine](https://ceoworld.biz). Since the meteoric rise of the smartphone and tech industries, many app developers have taken the opportunity to develop their own apps for different operating systems and made them available to download on the appleStore and/or the Google playstore. Today, there are over 2.2 billion iOS apps [Lifewire](https://lifewire.com) and 2.7 billion apps on the Google playstore [Statista](https://statista.com). On average, people spend on around 3 hours on their phone each day [Hacker Noon](https://hackernoon.com); with the majority of time being spent on different apps. Due to the huge popularity of apps, there is a wide variety of genres available for every type of audience with a wide range of pricing. Since many people invest so much time and effort into apps, app data has become an interesting and relevant topic to conduct more research on. 

This report focuses mainly on a sample of apps from the appleStore, where the composition of this sample will be analysed in depth. The report itself will contain 6 additional chapters: Chapter 2 looks at the source data and the general composition of the dataset. Chapter 3 compares the apps in the appleStore to that in the Google playstore; chapter 4 looks at the different app categories; chapter 5 analyses the pricing of the apps and looks at the relationship between app pricing and popularity. Finally chapter 6 is the report conclusion where we report our findings based on the analysis.

## Section 2: Description of Data

### 2.1 Source Data

The source data for our EDA comes from two datasets that our team discovered on [Kaggle](https://www.kaggle.com).  The first dataset contains information about apps in the Google Play Store.  The second contains information about apps in the Apple App Store.  First, we will describe the data pertaining to the Google Play Store.

#### Google Play Data Structure

The Google Play Store dataset contains records for 10,840 different apps.  The variables, and their descriptions are given below:

* **App** - The name of the application
* **Category** - The category that the app belongs in
* **Rating** - The current app rating on a scale from 1 to 5 (1 being the lowest)
* **Reviews** - The number of reviews for the app
* **Size** - Size of the app given in MB
* **Installs** - Number of installs given as categories (ex. 10,000+)
* **Type** - Tells which apps are free and which require payment
* **Price** - Gives the Price for each app
* **Content_Rating** - Indicates which age groups the app is approved for
* **Genres** - A further breakdown of the Category variable for each app
* **Last Updated** - Date the app was last updated
* **Current_Ver** - Current_Version of the app

```{r import Google Play Data, echo=F}
library("stringr")
# setwd("/Users/jonathangiguere/Desktop/DATS 6101/Projects/Project 1/Data-Science-Project-Final-master")
googlePlay <- read.csv('googleplaystore.csv')
googlePlay <- subset(googlePlay, select = -c(Android.Ver) )
googlePlay$Size <- str_remove(googlePlay$Size, 'M')
str(googlePlay)
```

### 2.2 Clean the Data

Check how many null values are in the playstore data frame:
```{r check NA, echo = F, include = T}

# review how many fields are NA
plotNA <- barplot(sapply(googlePlay,function(x)sum(is.na(x))),
                  col = "lightblue",
                  main = "Playstore NA Values",
                  ylab = "Count of Null Values",
                  xlab = "App Categories")
```


```{r clean NA, echo = F, include = T}
# create new data frame for apps where NA is null, then remove rating NAs from main playstore dataset
ratingNA <- subset(googlePlay, is.na(Rating))
googlePlay = na.omit(googlePlay)

# to double check that the NAs have been removed
sapply(googlePlay,function(x)sum(is.na(x)))
nrow(googlePlay)
```


The only field with null values is Rating with 1474 missing values. We will store these values in a separate data frame in case we would like to do future analysis on this data. Then we will remove them from the main playstore dataset.


Now we will check whether there are duplicates. We will do this by checking the difference between all and unique object.
```{r check dups, echo = F, include = T}

distinct <- nrow(googlePlay %>%
  distinct())

duplicates <- nrow(googlePlay) - distinct
duplicates
```
There are `r duplicates` duplicates in the data. The next step is to remove duplicates from playstore.
```{r remove dups, echo = F, include = T}

googlePlay=googlePlay[!duplicated(googlePlay), ]


# to double check that the duplicates have been removed
distinct <- nrow(googlePlay %>%
  distinct())

nrow(googlePlay) - distinct

```


#### Apple Store Data Structure

The Apple App Store dataset contains records for 7,197 different apps.  The variables, and their descriptions are given below:

* **App** - The name of the application
* **Category** - The category that the app belongs in
* **Rating** - The current app rating on a scale from 1 to 5 (1 being the lowest)
* **Reviews** - The number of reviews for the app
* **Size** - Size of the app given in MB
* **Price** - Gives the Price for each app
* **Content_Rating** - Indicates which age groups the app is approved for
* **Current_Ver** - Current_Version of the app

```{r import Apple Store Data, echo=F}
appleStore <- read.csv('applestore.csv')
appleStore <- subset(appleStore, select = -c(id, currency, rating_count_ver,
                                             user_rating_ver, sup_devices.num,
                                             ipadSc_urls.num, lang.num,
                                             vpp_lic))
colnames(appleStore) <- c('X', 'App', 'Size', 'Price', 'Reviews',
                          'Rating', 'Current_Ver', 'Content_Rating',
                          'Category')
appleStore$Size <- round(appleStore$Size/1000000, 0)
str(appleStore)
```

## Section 3: App Store Vs. Google Playstore

Even though the App Store is going to be analyzed in this report, a comparative analysis will be made before between the App Store and Google Playstore. This will serve as a guide to the structure of each platform and provide a visual comparison between the two.

The appleStore dataset contains 7197 different apps whereas the Google Playstore contains 10840.

```{r merge}

```

### 3.1 Rating Composition

```{r box1}
df1 <- data.frame(appleStore$Price)
df2 <- data.frame(googlePlay$Price)
ggplot(appleStore, aes(y = Price)) +
  geom_boxplot()
```

### 3.2 Rating Distribution of Apps

```{r hist}
hist(googlePlay$Rating, main = 'Histogram of frequency rating for the Google Playstore', xlab = 'Rating', col = viridis(20))
hist(appleStore$Rating, main = 'Histogram of frequency rating for the Apple Store', xlab = 'Rating', col = viridis(12))
```

### 3.3 Price Composition

```{r box2}
```

### 3.4 Free Vs. Paid

```{r pie}
#appleStore$Type <- factor(ifelse(appleStore$Price == 0.00, 0, 1), labels = c('Free', 'Paid'))

#ggplot(appleStore, aes(x = " ", y=Type, fill = Type)) +
  #geom_bar(stat = "identity") +
  #ggtitle('Price Composition of Apps in the Appstore') +
  #coord_polar("y", start=0) +
  #theme_void()

#ggplot(googlePlay, aes(x = " ", y=Type, fill = Type)) +
  #geom_bar(stat = "identity") +
  #ggtitle('Price Composition of Apps in the Google Playstore') +
  #coord_polar("y", start=0) +
  #theme_void()
```

## Section 4: App Categories

### 4.1 SMART Question

Does app category have a correlation with rating and what app categories in the Googleplay have the highest priced apps?

### 4.2 Category Analysis

By grouping apps by category, we can determine the average rating by category and average price, as well as overall number of apps and number of apps  by category. The most common categories are family, game and tools; below the category distrubution of apps is shown.

```{r bar2, echo = F}
ggplot(googlePlay, aes(Category)) + 
        geom_bar(fill = viridis(33), color = 'black') +
        ggtitle('Number of Apps in the Google Playstore by Category') +
        ylab('Number of Apps') +
        coord_cartesian(ylim = c(0,1250)) +
        theme_classic() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))
```

The highest rated categories on average are Events, Education and Art & Desgin, while the lowest rated categories are Maps & Navigation, Tools and Dating.

```{r bary, echo = F}
catrating <- aggregate(googlePlay$Rating, list(googlePlay$Category), mean, na.rm = TRUE)
names(catrating) <- c('Category', 'Avg_Rating')

ggplot(catrating, aes(x = Category, y = Avg_Rating)) +
        geom_bar(stat = 'Identity', fill = viridis(33), color = 'black') +
        ggtitle('Average Rating of Apps in the Google Playstore by Genre') +
        ylab('Average Rating') +
        coord_cartesian(ylim = c(3.75, 4.5)) +
        theme_classic() +
        theme(plot.title = element_text(hjust = 0.5)) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))
```

Highest Average Rating by Category

```{r top rating}
head(catrating[order(catrating$Avg_Rating, decreasing = TRUE),], n = 5)
```


Lowest Average Rating by Category
```{r bottom rating}
tail(catrating[order(catrating$Avg_Rating, decreasing = TRUE),], n = 5)
```

``` {r type2, include = F}
free <- subset(googlePlay, googlePlay$Type == 'Free')
paid <- subset(googlePlay, googlePlay$Type == 'Paid')

print(nrow(free))
print(nrow(paid))
```

Even though most of the apps in the Googleplaystore are free; with `r nrow(free)` free apps and `r nrow(paid)` apps. Out of those paid apps, the most expensive ones on average are Finance and Lifestyle apps by a long shot. 

```{r barx, echo = F}
paid$Cost = as.numeric(gsub("\\$", "", paid$Price))
catprice <- aggregate(paid$Cost, list(paid$Category), mean)
names(catprice) <- c('Category', 'Avg_Price')

ggplot(catprice, aes(x = Category, y = Avg_Price)) +
        geom_bar(stat = 'Identity', fill = viridis(28), color = 'black') + 
        ggtitle('Average Price of Paid Apps in the Google Playstore by Category') +
        ylab('Average Price ($)') +
        coord_cartesian(ylim = c(0,20)) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Highes Average Price by Category

```{r top paid}
head(catprice[order(catprice$Avg_Price, decreasing = TRUE),], n = 5)
```

### 4.3 Which app category in the Googleplaystore receives most high ratings?

In this section, we would like to identify the most well rated app categories in the Googleplaystore. We define ‘most well rated’ as ratings that are greater or equal to 4.8. The score 4.8 is chosen because 95% of app ratings in the Googleplaystore are below 4.8. In other words, only 5% of the Googleplaystore apps have ratings that are greater or equal to 4.8. Therefore, the 95th percentile helps us to filter the best rated apps.

```{r hist2}
hist(googlePlay$Rating, main = 'Histogram of frequency rating for the Google Playstore', xlab = 'Rating', col = viridis(20))
```

```{r percentile}
quantile(googlePlay$Rating, .95)
```

```{r best rated}
top_rated <- subset(googlePlay, googlePlay$Rating >= 4.8)
top_count <- plyr::count(top_rated$Category)
names(top_count) <- c('Category', 'Count')
top_count <- top_count[order(top_count$Count, decreasing = TRUE),]
```

To do this, a new subset was created from the Googleplaystore dataframe which only contained apps with a rating of 4.8 or greater. There are `r nrow(top_rated)`apps which fall within the highly rated threshold of 4.8. After this, the number of apps from each catergory within this subset was done using the count function. This count was then plotted into a bar chart to graphically demonstrate the distrubution of high ratings within different categories.

```{r topplot}
ggplot(top_count, aes(x = reorder(Category, -Count), y = Count)) +
        geom_bar(stat = 'Identity', fill = plasma(32), color = 'black') + 
        ggtitle('Highly Rated Apps in the Google Playstore by Category') +
        ylab('Total') +
        coord_cartesian(ylim = c(0,60)) +
        theme_bw() +
        theme(plot.title = element_text(hjust = 0.5)) + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 7, color = 'black')) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Catergories with the highest ratings are Family (by a large margin), then Medical, Lifestyle and Health & Fitness being very similar.
```{r topr}
head(top_count[order(top_count$Count, decreasing = TRUE),], n = 5)
```

Overall; the highest rated apps on average are in the Events category, the most expensive apps are in the Finance category and the category with the most highly rated apps are Family.

## Section 5: App Pricing

KEY QUESTION: How does the price influence the distribution of apps across category, rating and content rating

``` {r type, echo = F}
# Create two new dataframes for free and paid apps separately
freeApps <- subset(googlePlay, googlePlay$Price == 0.00)
paidApps <- subset(googlePlay, googlePlay$Price != 0.00)

nrow(freeApps)
nrow(paidApps)
```

There are `r nrow(freeApps)` free apps and `r nrow(paidApps)` paid apps corresponding to `r format((nrow(freeApps)/nrow(googlePlay))*100, digits=3)`% and `r format((nrow(paidApps)/nrow(googlePlay))*100, digits=3)`% respectively in Google playstore 


How do users rate free and paid apps across different categories
```{r echo=F}
# Create new dataframe and new column, assign categorical values of 0 and 1 for free and paid apps respectively   
  googlePlay_cat <- googlePlay[which(googlePlay$Price == 0.00 |googlePlay$Price != 0.00), ]
  googlePlay_cat$PriceCat <- factor(ifelse(googlePlay_cat$Price == 0.00, 0, 1), labels = c("Free", "Paid"))
  ggplot(googlePlay_cat, aes(x = Category, y = Rating)) +
    geom_boxplot(fill=viridis(61)) +
    #ggtitle("Boxplot of User Rating by Genre across Free Vs Paid Apps")
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6, color = 'black')) +
    facet_grid(. ~ PriceCat)
```
What kind of ratings does the different content rating categories have for both free and paid apps
```{r }
  ggplot(googlePlay_cat, aes(x = Content.Rating, y = Rating)) +
      geom_boxplot(fill=viridis(10)) +
      ggtitle("Boxplot of User Rating by Content Rating across Free Vs Paid Apps")+
      facet_grid(. ~ PriceCat) +
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5)) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6, color = 'black'))      
```

How are free apps distributed by category
```{r d, echo = F}
# create a dataframe sorted by count of category for free apps and plot
freeAppsCount <- freeApps %>% group_by(Category) %>% dplyr::summarise(NumOfApps = n()) %>% arrange(desc(NumOfApps)) %>% mutate(Category = factor(Category, levels = unique(Category)))
freeAppsCount
 ggplot(freeAppsCount, aes(x=Category, y=NumOfApps)) +
  geom_bar( fill = viridis(33), color = 'black', stat = 'identity') +
  ggtitle('Free Apps in the Google Playstore by Category') + 
  xlab('Category') + ylab('Number of Apps') + 
  #coord_cartesian(ylim = c(0, 1800)) + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))

```

Top 5 Category for Free Apps

```{r }
  head(freeAppsCount, n = 5)
```
```{r echo = F}
# create a dataframe sorted by count of category for paid apps and plot

  paidAppsCount <- paidApps %>% group_by(Category) %>% dplyr::summarise(NumOfApps = n()) %>% arrange(desc(NumOfApps)) %>% mutate(Category = factor(Category, levels = unique(Category)))
ggplot(paidAppsCount, aes(x = Category, y = NumOfApps)) +
  geom_bar(fill = viridis(28), color = 'black', stat = 'identity') +
  ggtitle('Paid Apps in Google Playstore by Category') + 
  xlab('Category') + ylab('Number of Apps') + 
  coord_cartesian(ylim = c(0, 200)) + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))
```

Top 5 Category for Paid Apps
```{r }
  head(paidAppsCount, n = 5)
```
In the Playstore, apps are either free or paid. 
What genre of apps are rated more in the free vs paid apps?

```{r echo = F}
# create new dataframe containing different categories and their respective average rating for free apps
catFreeRating = freeApps %>% group_by(Category) %>% dplyr::summarise( Avg_rating = mean(Rating)) %>% 
arrange(desc(Avg_rating)) %>% mutate(Category = factor(Category, levels = unique(Category)))

ggplot(catFreeRating, aes(x = Category, y = Avg_rating)) + 
  geom_bar(stat = 'Identity', fill = viridis(33), color = "black") +
  ggtitle("Average Rating of Free Apps by Category") +
   ylab('Average Rating') +
  coord_cartesian(ylim = c(1.5, 5.5)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9, color = 'black')) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```


```{r echo = F}
# create new dataframe containing different categories and their respective average rating for paid apps
#catPaidRating = aggregate(paidApps$Rating, list(paidApps$Category), mean)
#names(catPaidRating) = c("Category", "Avg_rating")

catPaidRating = paidApps %>% group_by(Category) %>% dplyr::summarise( Avg_rating = mean(Rating)) %>% 
arrange(desc(Avg_rating)) %>% mutate(Category = factor(Category, levels = unique(Category)))

ggplot(catPaidRating, aes(x = Category, y = Avg_rating)) + 
  geom_bar(stat = 'Identity', fill = viridis(28), color = "black") +
  ggtitle("Average Rating of Paid Apps by Category") +
   ylab('Average Rating') +
  coord_cartesian(ylim = c(1.5, 5.5)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 9, color = 'black')) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

Do users rate free apps better than paid apps?
Top 5 rated category in Free Apps 
```{r }

  head(catFreeRating, n = 10)
```

Top 5 rated category in Paid Apps 
```{r }

  head(catPaidRating, n = 10)
```

```{r echo = F}
ggplot(freeApps, aes(Content.Rating)) +
  geom_bar(fill = viridis(6), color = 'black') +
  ggtitle('Free Apps in Google Playstore by Content Rating') + 
  xlab('Content Rating') + ylab('Number of Apps') + 
  coord_cartesian(ylim = c(0, 10000)) + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))
```

```{r echo = F}
ggplot(paidApps, aes(Content.Rating)) +
  geom_bar(fill = viridis(4), color = 'black') +
  ggtitle('Paid Apps in Google Playstore by Content Rating') + 
  xlab('User Rating') + ylab('Number of Apps') + 
  coord_cartesian(ylim = c(0, 1000)) + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8, color = 'black'))
```